PyQBench: a Python library for benchmarking gate-based quantum computers

### 2022-12-12 txt/manuscript.tex
Update title and corresponding author
Update manuscript
creating appendix and rewriting mathemaytical prelimaties and parametrized family

%% 
%% Copyright 2007, 2008, 2009 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01

@@

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}
\usepackage{mathtools}
%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
\usepackage{lineno}

@@

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
%\DeclareMathOperator{\diag}{diag}
%\DeclareMathOperator{\diag}{diag}
\journal{SoftwareX}

@@

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\title{}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \address[label1]{}
%% \address[label2]{}

\author{Konrad Jałowiecki}
\author{Paulina Lewandowska}
\author{Profesor Kierownik \L ukasz Pawela}

@@

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

@@

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{AB}}$ on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{AB}}$ and measures the system $A$.
	\item Alice performs the conditional binary measurement $\PP_{V_i}$ on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures the system $B$ of the input state  $\ket{\psi_{AB}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{AB}}$ and 
receives an output label $i$.
\item  
Based on Holevo--Helstrom theorem, Bob performs a conditional binary 
measurement	$\PP_{V_j^\dagger}$ on part $B$.
\item To calculate the probability of correct discrimination, they include just those cases for which $i = j$.
\item Bob measures part $B$ of the input state  $\ket{\psi_{AB}}$ and makes a
decision based on received label $j$. If $i=j=0$, then he decides that the 
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

@@

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U}, \mathcal{P}_{V_k}, i, j) \\ 
(\Phi_{\Id}, \mathcal{P}_{V_k}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
		\begin{equation}
		p = \frac{\#(\Phi_U, \PP_{V_i},i,0) + \#(\Phi_\Id, \PP_{V_i},i,1)}{\#(\Phi_U, \PP_{V_i},i,j) + \#(\Phi_\Id, \PP_{V_i},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{AB}}$ on 
systems $A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U}$,  on part $A$ of the input state  $\ket{\psi_{AB}}$.
\item Alice and Bob implement controlled unitary $V_0 \oplus V_1$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$.
Whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

 We will also need a linear mapping transforming $M_{d_1}$ into
$M_{d_2}$, which will be denoted 
\begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
There
exists a bijection between introduced linear mappings $\Phi$ and set of matrices $M_{d_1d_2}$,  known as the Choi-
Jamio{\l}kowski isomorphism. 
For a given linear mapping $\Phi_M$ corresponding  Choi operator $\mathcal{J}(\Phi)$ is explicitly written as 
\begin{equation}
\mathcal{J}(\Phi) \coloneqq \sum_{i,j=1}^{d- 1} \Phi(\ketbra{i}{j}) \otimes \ketbra{i}{j}. \end{equation}
Finally, we introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
In this work we will consider a special class of quantum channels, called unitary channels. Let $U \in \UU_d$ be a unitary operator. A unitary channel
$\Phi_{U}$ is defined as $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$. A general quantum
measurement, that is a positive operator valued measure (POVM) $\PP$ is a
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
\emph{effects}, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all the effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. Every von Neumann measurement can be
parameterized by a unitary matrix and hence we will use the notation $\PP_{U}$
for a von Neumann measurement with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i+1$-th column of the unitary matrix $U$. The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$

@@

Let us  consider a  linear mapping transforming  square matrices into square 
matrices that is $\Phi: M_{d_1} \to M_{d_2}$. 
We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}

The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum operations, $\TT$  and $\SS$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\TT, \SS) =  \frac12 + \frac14 \| \TT - \SS \|_\diamond.
\end{equation}
In the case of discrimination between two von Neumann measurements $\PP_U$ and 
$\PP_\Id$ we know that (from Theorem 1 in \cite{puchala2018strategies})
\begin{equation}
\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
\Phi_\Id\|_\diamond,
\end{equation}
where $\Phi_U$ is unitary channel and $\diaguni_d$ be the set of diagonal 
unitary matrices of dimension $d$.

@@

Before we proceed to presenting our main results, we need to briefly discuss the
problem of discrimination of unitary channels.  For this purpose, we introduce the notion of numerical range of
a given matrix $A \in M_d$. The set \begin{equation}
W(A) =\{\bra{x}A\ket{x}: \ket{x} \in 
\mathbb{C}^d, \;
\;\braket{x}{x}=1\}.
\end{equation}   
is the numerical range of $A$.
The detailed properties of the numerical range and its generalizations we can see on the website~\cite{nr}.

 Let us now recall the well-known result for the
distinguishability of unitary channels.
	Let $U \in \UU_d$ and $\Phi_U$ be a unitary 
	channel. 
	Then, the following equation holds
	\begin{equation}
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
	\end{equation}
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. And hence the optimal probability to discriminate two unitary channel is given by 
	\begin{equation}
	p_{\text{succ}}(\Phi_U, \Phi_\Id) = \frac{1}{2} + \frac{1}{2} \sqrt{1-\nu^2}.
	\end{equation}

@@

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H 
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger
\end{equation}
%\begin{equation}
%U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
%\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we present theoretical probability of correct 
discrimination between these measurements. Moreover, we create the optimal 
theoretical strategy of their discrimination. 
The otimal probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the proper input state, (see sec.~\ref{sec:example_discriminator}) 
and final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

In this section we calculate the theoretical probability of correct discrimination between two von Neumann measurements $\PP_\1$ and $\PP_U$ for $U = H \diag(1, e^{i \phi}) H^\dagger$, where  $\phi \in [0, 2\pi)$.  To do that, we will present an auxiliary lemma.   
\begin{lemma}\label{lemma:min-e-optimal}
Let $U = H \diag(1, e^{i \phi}) H^\dagger$, $\phi \in [0, 2\pi)$ and	let 
$\Phi_U$ and $\Phi_\Id$ be two unitary channels. Then, the following equation holds 
\begin{equation}
\min_{E \in \diaguni_2} \|\Phi_{UE} - 
\Phi_\Id\|_\diamond = \|\Phi_{U} - 
\Phi_\Id\|_\diamond,
\end{equation}
\end{lemma}

%	For simplify notation, let 	$ \nu \coloneqq \max_{E \in \diaguni_2 } \nu_{E} $.
 The celebrated Hausdorf-T{\o}plitz theorem~\cite{hausdorff, toeplitz} states that
$W(A)$ of any matrix $A \in M_d$ is a convex set and therefore
\begin{equation}
W(A) = \{ \tr(A \rho): \rho \in \Omega_d\}. 
\end{equation}
So we obtain that 
	\begin{equation}
	\min_{\ket{x} \in \mathbb{C}^2:   \proj{x} = 1} |\bra{x}U^\dagger\ket{x}| = 
	\min_{\rho \in \Omega_2} |\tr(U^\dagger\rho)|. 
	\end{equation}
	And then we have 
	 \begin{equation}
 \max_{E \in \diaguni_2 } \nu_{E}  = \max_{E \in \diaguni_2 }  \min_{\rho \in 
		\Omega_2} \left| \tr \left( \rho U E \right) \right|
	\end{equation}
	For that, our task is reduce to show that
	\begin{equation}
	\forall_{E \in \diaguni_2} \,\, | \tr \left(\rho U E\right) | \le \mu
	\end{equation}
	where $\mu \coloneqq \max_{E \in \diaguni_2 } \nu_{E}.$

	Let us define $E = \left(\begin{array}{cc}E_0&0\\0&E_1\end{array}\right)  $ 
	and take $\rho = 
	\left(\begin{array}{cc}\frac{1}{2}&0\\0&\frac{1}{2}\end{array}\right) $. 
	From spectral theorem, let us decompose $U$ as
	\begin{equation}
	U= \lambda_1 \ketbra{x_1}{x_1} + \lambda_2 \ketbra{x_1}{x_2}, 
	\end{equation}
	where  for eigenvalue $\lambda_1 = 1$, the corresponding 
	eigenvector is 
	of the form $\ket{x_1} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\\frac{1}{\sqrt{2}}\end{array}\right]
	$,
	whereas for  $\lambda_2 = e^{i \phi}$ we have $\ket{x_2} = \left[\begin{array}{c}\frac{1}{\sqrt{2}}\\-\frac{1}{\sqrt{2}}\end{array}\right] 
	$.
	Then, we have 
	\begin{equation}
	\begin{split}
	& \forall E \in \diaguni_2 \,\,\, | \tr (\rho U E) | = \frac{1}{2}  \left| \tr \left(
	H \diag(1, e^{i\phi}) H^\dagger E \right) \right| =  \\ &
	\frac{1}{2} \left| \tr\left((   \proj{x_1} +e^{i \phi}\proj{x_2} ) E \right) 
	\right|  = 
	\frac{1}{2} \left|  \bra{x_1} E \ket{x_1} +  e^{i \phi}\bra{x_2} E \ket{x_2} 
	\right| = \\& 
	\frac{1}{2} \left| \frac{E_0 + E_1}{2} + e^{i \phi } \frac{E_0+E_1}{2} \right| 
	= 
	\frac{\left| 1+ e^{i \phi } \right|}{2} \left| \frac{E_0 + E_1}{2} \right| \le 
	\mu, 
	\end{split}
	\end{equation}
	which completes the proof.
\end{proof}
\begin{theorem}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ for $U = H \diag(1, e^{i \phi}) H^\dagger$,
where $\phi \in [0, 2\pi)$ is given by
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}

\begin{proof}
	From Holevo-Helstrom theorem we obtain
	\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{1}{4} \| \PP_{U} - \PP_{\Id} \|_\diamond.
	\end{equation}
	From Theorem 1 in~\cite{puchala2018strategies} we have 
	\begin{equation}
	\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond. 
	\end{equation}
	From Lemma~\ref{lemma:min-e-optimal} we know that for 
	$U =  H \diag(1, e^{i \phi}) H^\dagger$ it also holds that
	\begin{equation}
		\min_{E \in \diaguni_2} \|\Phi_{UE} - 
	\Phi_\Id\|_\diamond = \|\Phi_{U} - 
	\Phi_\Id\|_\diamond,
	\end{equation} which is exactly equal to 
	\begin{equation}
	\|\Phi_{U} - 
	\Phi_\Id\|_\diamond = 2\sqrt{1 - \nu^2} = |1-e^{i   \phi }|. 
	\end{equation}
%	Finally, we obtain that
%	\begin{equation}
%	\| \PP_{U} - \PP_{\Id} \|_\diamond =  |1-e^{i   \phi }|.
%	\end{equation}
	It implies that
	\begin{equation}
	p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1-e^{i \phi}|}{4},
	\end{equation} which completes the proof.
\end{proof}

\subsection{Realization}\label{sec:example_realization}
Now we focus on the realization of the discrimination protocol. We will 
discriminate between $\PP_U$ and $\PP_\1$. To do so, we need to know exact 
forms of the input state and final measurement. We will indicate the optimal 
input state in subsection~\ref{sec:example_discriminator} and the final 
measurement in subsection~\ref{sec_example_final_measurement}.

@@

\begin{proposition}\label{prop-discrim}
Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
and $\PP_\1$, $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
2\pi)$.  The optimal input state has the form
\begin{equation}
\ket{\psi_{AB}} = \frac{1}{\sqrt{2}} |\Id_2 \rangle \rangle.
\end{equation}
\end{proposition}

>>>>

@@

\usepackage{amssymb}

\usepackage{amsthm}
\usepackage{mathtools}

\usepackage{lineno}

@@

\newcommand{\eg}{{\emph{e.g.\/}}}
\newcommand{\ie}{{\emph{i.e.\/}}}
\newcommand{\ket}[1]{\ensuremath{|#1\rangle}}
\newcommand{\bra}[1]{\ensuremath{\langle#1|}}
\newcommand{\ketbra}[2]{\ensuremath{\ket{#1}\bra{#2}}}
\newcommand{\proj}[1]{\ensuremath{\ketbra{#1}{#1}}}
\newcommand{\braket}[2]{\ensuremath{\langle{#1}|{#2}\rangle}}
\newcommand{\floor}[1]{\ensuremath{\lfloor #1 \rfloor}}
\newcommand{\complexity}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\new}[1]{ \textcolor{red}{#1} }
\newcommand{\1}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\Id}{{\rm 1\hspace{-0.9mm}l}}
\newcommand{\connected}{\sim}
\newcommand{\SPAN}{\mathrm{span}}
\newcommand{\Lrm}{\ensuremath{\mathrm{L}}}
\newcommand{\Urm}{\ensuremath{\mathrm{U}}}
\newcommand{\ee}{\ensuremath{\mathrm{e}}}
\newcommand{\dd}{\ensuremath{\mathrm{d}}}
\newcommand{\ii}{\ensuremath{\mathrm{i}}}
\newcommand{\EE}{\mathcal{E}}
\newcommand{\XX}{\mathcal{X}}
\newcommand{\MM}{\mathcal{M}}
\newcommand{\NN}{\mathcal{N}}
\newcommand{\DD}{\mathcal{D}}
\newcommand{\TT}{\mathcal{T}}
\newcommand{\PP}{\mathcal{P}}
\renewcommand{\SS}{\mathcal{S}}
\newcommand{\UU}{\mathcal{U}}
\newcommand{\HH}{\mathcal{H}}
\newcommand{\DU}{\mathcal{DU}}
\newcommand{\NOT}{\sigma_x}
\newcommand{\idop}[1][\XX]{\ensuremath{\1_{#1}}}
\newcommand{\diaguni}{\ensuremath{\mathcal{DU}}}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{tr}}
\journal{SoftwareX}

@@

\title{PyQBench: a Python library for benchmarking gate-based quantum computers}

\author{Konrad Jałowiecki\corref{cor1}}
\ead{dexter2206@gmail.com}
\cortext[cor1]{Corresponding author}

\author{Paulina Lewandowska}
\author{Profesor Kierownik \L ukasz Pawela}

@@

%\PACS 03.67.−a \sep 03.67.Lx

\MSC 81P68

@@

\begin{enumerate}
	\item Alice and Bob share an entangled state $\ket{\psi_{0}}$ (discriminator) on systems 
	$A$ and $B$.
	\item Alice performs one of two known measurements (either $\PP_{\Id}$ or 
	$\PP_{U}$) on part $A$ of the input state  $\ket{\psi_{0}}$ and measures the system $A$.
	\item Alice performs the conditional binary measurement $\PP_{V_i}$ on the part 
	$B$ based on the output label $i$. 
	\item  Bob measures the system $B$ of the input state  $\ket{\psi_{0}}$ and 
	makes a decision based on the received label $j$. If $j=0$, then he decides 
	that the the black box contains $\PP_U$. Otherwise, he decides that the 
	black box contains $\PP_{\Id}$. 
\end{enumerate}

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{0}}$ (discriminator) on systems 
$A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or 
$\Phi_{U^\dagger}$,  on part $A$ of the input state  $\ket{\psi_{0}}$.
\item Alice measures part $A$ of the input state  $\ket{\psi_{0}}$ and 
receives an output label $i$.
\item  
Based on Holevo--Helstrom theorem, Bob performs a conditional binary 
measurement	$\PP_{V_j}$ and implements the unitary channel 	$\Phi_{V_j^\dagger}$ on part $B$.
\item To calculate the probability of correct discrimination, they include just those cases for which $i = j$.
\item Bob measures part $B$ of the input state  $\ket{\psi_{0}}$ and makes a
decision based on received label $j$. If $i=j=0$, then he decides that the 
black box contains $\PP_U$. Otherwise, he decides that the black box contains
$\PP_{\Id}$.
\end{enumerate}

@@

Due to the fact that the measurement result is nondeterministic, we are forced to consider the following circuits
in the postselection procedure
\begin{equation}
\begin{split}
(\Phi_{U^\dagger}, \Phi_{V_k^\dagger}, i, j) \\ 
(\Phi_{\Id}, \Phi_{V_k^\dagger}, i, j)
\end{split}
\end{equation}
where $k=\{0,1\}, i=\{0,1\}$ and $ j=\{0,1\}$. 
The optimal probability of correct discrimination between $\PP_{U} $ and $\PP_\Id$ is as follows.
		\begin{equation}
		p = \frac{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,0) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,1)}{\#(\Phi_{U^\dagger}, \Phi_{V_i^\dagger},i,j) + \#(\Phi_\Id, \Phi_{V_i^\dagger},i,j)}, 
		\end{equation}
where $i=\{0,1\}$ and $ j=\{0,1\}$.

@@

\begin{enumerate}
\item Alice and Bob prepare an entangled state $\ket{\psi_{0}}$ on 
systems $A$ and $B$.
\item Alice performs one of two quantum channels, either $\Phi_{\Id}$ or
$\Phi_{U^\dagger}$,  on part $A$ of the input state  $\ket{\psi_{0}}$.
\item Alice and Bob implement  direct sum of unitary channels $\Phi_{V_0^\dagger} \oplus \Phi_{V_1^\dagger}$ obtained from 
Holevo--Helstrom theorem.	
\item Alice and Bob prepare the measurement $\Delta$ in computational basis on 
their systems.
\item Bob makes a decision based on received label $j$. If $j=0$, then he
decides that the the black box contains $\PP_U$. Otherwise, he decides that the
black box contains $\PP_{\Id}$.
\end{enumerate}

@@

Let $M_{d_1,d_2}$ be the set of all matrices of dimension $d_1 \times d_2$ over
the field $\mathbb{C}$. For  simplicity, square matrices will be denoted by
$M_d$.  The subset of $M_d$ consisting of Hermitian matrices of dimension $d$
will  be  denoted  by $\HH_d$,  while  the  set  of  positive semidefinite
matrices of dimension $d$ by $\HH_d^+$. The set of quantum states, that is
positive semidefinite operators having trace equal to one, will be denoted
$\Omega_d$. 
 The subset of $M_d$ consisting of unitary matrices will be denoted
by $\UU_d$, while its subgroup of diagonal unitary operators will be denoted by
$\DD \UU_d$. 
	An operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a stochastic matrix if $a_{i,j} \ge 0$   and $\sum_{i} a_{i,j} = 1$, whereas an operator $A = \left( a_{i,j}\right)_{i,j} \in M_d$ is said to be a double stochastic matrix if $A$ is a stochastic matrix and $\sum_{j} a_{i,j} = 1$.

 We will also need a linear mapping transforming $M_{d_1}$ into
$M_{d_2}$, which will be denoted 
\begin{equation}
\Phi: M_{d_1 } \rightarrow M_{d_2}.
\end{equation} 
There
exists a bijection between the set of linear mappings $\Phi$ and the set of matrices $M_{d_1d_2}$,  known as the Choi-Jamio{\l}kowski isomorphism. 
For a given linear mapping $\Phi$ the corresponding Choi operator $\mathcal{J}(\Phi)$ is explicitly written as 
\begin{equation}
\mathcal{J}(\Phi) \coloneqq \sum_{i,j=0}^{d- 1} \Phi(\ketbra{i}{j}) \otimes \ketbra{i}{j}. \end{equation}
We introduce a special subset of all mappings $\Phi$, called quantum channels, which are completely positive
and trace preserving (CPTP).
In this work we will consider a special class of quantum channels, called unitary channels.  A 
channel
$\Phi_{U}$ is said to be a unitary channel if it has the following form $\Phi_U(\cdot) = U \cdot U^\dagger$ for  $U \in 
\UU_d$.

Finally, we define a quantum
measurement, that is a positive operator valued measure (POVM) denoted by $\PP, \mathcal{Q}$ etc. This is a 
collection of positive semidefinite operators $\{E_1, \ldots, E_m \}$ called
effects, which sum up to identity, \ie $ \, \, \sum_{i=1}^m E_i = \1$. If
all effects are rank-one projection operators, then such a measurement is
called a von Neumann measurement. 
Every von Neumann measurement can be
parameterized by a unitary matrix $U$ with effects $\{\proj{u_0}, \ldots, \proj{u_{d-1}}\}$,
where $\ket{u_i}$ is the $i$-th column of the unitary matrix $U$.  Hence we will use the notation $\PP_{U}$.  The action of
quantum measurement $\PP_{U}$ on some state $\rho \in \Omega_d$ can be
seen as  a measure-and-prepare quantum channel as follows 
$
\PP_{U} : \rho \rightarrow \sum_{i=0}^{d-1} \bra{u_i} \rho \ket{u_i} \proj{i}.
$ Moreover, observe that each von Neumann measuement $\PP_{U}$ can be written as compostion of a unitary channel $\Phi_{U^\dagger}$ and the maximally dephasing channel $\Delta$, that means $\PP_{U} = \Delta \circ \Phi_{U^\dagger}$.

@@

Let us  consider a  linear mapping transforming  square matrices into square 
matrices that is $\Phi: M_{d_1} \to M_{d_2}$. 
We define the diamond norm of a map $\Phi$ as 
\begin{equation}
\|\Phi\|_\diamond = \max_{\|X\|_1=1} \| \left(\Phi \otimes \1\right) (X) \|_1.
\end{equation}
The celebrated result by Helstrom~\cite{helstrom1976quantum} gives the optimal  probability of correct discrimination between two quantum operations, $\Psi$  and $\Phi$, 
in terms of their distance with the use of the diamond norm
\begin{equation}
p_{\text{succ}}(\Psi, \Phi) =  \frac12 + \frac14 \| \Psi - \Phi \|_\diamond.
\end{equation}
from Theorem 1 in \cite{puchala2018strategies}, the distance between two von Neumann measurements $\PP_U$ and 
$\PP_\Id$ is given by
\begin{equation}
\|\PP_U - \PP_\Id\|_\diamond = \min_{E \in \diaguni_d} \|\Phi_{UE} - 
\Phi_\Id\|_\diamond,
\end{equation}
where $\Phi_U$ is unitary channel and $\diaguni_d$ be the set of diagonal 
unitary matrices of dimension $d$.

@@

Before we proceed to presenting our main results, we need to briefly discuss the
problem of discrimination of unitary channels.  For this purpose, we introduce the notion of numerical range of
a given matrix $A \in M_d$. The set \begin{equation}
W(A) =\{\bra{x}A\ket{x}: \ket{x} \in 
\mathbb{C}^d, \;
\;\braket{x}{x}=1\}.
\end{equation}   
is called numerical range of $A$.
The detailed properties of the numerical range and its generalizations we can see on the website~\cite{nr}.

	Let $U \in \UU_d$ and $\Phi_U$ be a unitary 
	channel. 
	Then, the following equation holds
	\begin{equation}
	\| \Phi_U  - \Phi_{\1} \|_\diamond = 2 \sqrt{1-\nu^2},
	\end{equation}
	where $\nu = \min_{x \in W(U^\dagger)} |x|  $. Finally, the optimal probability to discriminate two unitary channel is given by 
	\begin{equation}
	p_{\text{succ}}(\Phi_U, \Phi_\Id) = \frac{1}{2} + \frac{1}{2} \sqrt{1-\nu^2}.
	\end{equation}

@@

Let us focus on single-qubit von Neumann measurements $\PP_\1$ and $\PP_U$.
Assume that the unitary matrix $U$ is of the form 
\begin{equation}
U = H 
\left(\begin{array}{cc}1&0\\0&e^{i \phi}\end{array}\right)  H^\dagger
\end{equation}
%\begin{equation}
%U = H \diag (1, \ee^{\ii \phi}) H^\dagger,
%\end{equation}
where $H$ is the Hadamard matrix of dimension two and $\phi \in [0, 2 \pi)$.
In this section we present theoretical probability of correct 
discrimination between these measurements. Moreover, we create the optimal 
theoretical strategy of their discrimination. 
The otimal probability will be calculated in 
subsection~\ref{sec:example_theoretical_probability}. Later, we will 
discuss the realization of this scheme in 
subsection~\ref{sec:example_realization}. For such realization, we 
will need the optimal input state (see sec.~\ref{sec:example_discriminator}), 
and the optimal final measurement (see sec.~\ref{sec_example_final_measurement}).

@@

\begin{theorem}\label{th:probability}
The optimal probability of correct discrimination between von Neumann
measurements $\PP_U$ and $\PP_{\Id}$ for $U = H \diag(1, e^{i \phi}) H^\dagger$,
where $\phi \in [0, 2\pi)$ is given by
\begin{equation}
p_{\text{succ}}(\PP_{U}, \PP_{\Id}) = \frac{1}{2} + \frac{|1 - e^{i \phi}  |}{4} . 
\end{equation}
\end{theorem}

\subsection{Realization}\label{sec:example_realization}

To indicate the optimal strategy  we will present two propositions. The first one is concentrated around the discriminator as the optimal input state of discrimination strategy, whereas the second one describes the optimal final measurement.

@@

\begin{proposition}\label{prop-discrim}
Consider the problem of discrimination between von Neumann measurements $\PP_U$ 
and $\PP_\1$, $U = H\diag(1, e^{i \phi}) H^\dagger $ and $\phi \in [0, 
2\pi)$.  The optimal input state has the form
\begin{equation}
\ket{\psi_{0}} = \frac{1}{\sqrt{2}} |\Id_2 \rangle \rangle.
\end{equation}
\end{proposition}

\subsubsection{Conditional measurement}\label{sec_example_final_measurement}
